---
title: "ML Project"
author: "Chris Trinh"
date: "Saturday, February 21, 2015"
output: html_document
---

# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 


# Load the Data and Explore
```{r warning=FALSE}
library(knitr)
library(lattice)
library(ggplot2)
library(caret)
library(randomForest)


set.seed(100)
opts_chunk$set(echo=TRUE,results="show",fig.align="center",cache=TRUE)

#read data
setwd("C:/Users/Chris/Documents")
train <- read.csv("./pml-training.csv")
test <- read.csv("./pml-testing.csv")
```

# Look at the Data and Build Training and Validation Sets
```{r}
dim(train); dim(test)

table(train$classe)

set.seed(123456)
trainset <- createDataPartition(train$classe, p = 0.8, list = FALSE)
Training <- train[trainset, ]
Validation <- train[-trainset, ]
```


# Selection
```{r}
# exclude near zero variance features
nzvcol <- nearZeroVar(Training)
Training <- Training[, -nzvcol]

# exclude columns with m40% ore more missing values exclude descriptive
# columns like name etc
cntlength <- sapply(Training, function(x) {
  sum(!(is.na(x) | x == ""))
})
nullcol <- names(cntlength[cntlength < 0.6 * length(Training$classe)])
descriptcol <- c("X", "user_name", "raw_timestamp_part_1", "raw_timestamp_part_2", 
                 "cvtd_timestamp", "new_window", "num_window")
excludecols <- c(descriptcol, nullcol)
Training <- Training[, !names(Training) %in% excludecols]
```


#  Train the Model
```{r}
rfModel <- randomForest(classe ~ ., data = Training, importance = TRUE, ntrees = 10)
```


# Check Training Accuracy
```{r}
ptraining <- predict(rfModel, Training)
print(confusionMatrix(ptraining, Training$classe))
```


# Validate Accuracy
```{r}
pvalidation <- predict(rfModel, Validation)
print(confusionMatrix(pvalidation, Validation$classe))
```


# Predict the Test Set
```{r}
ptest <- predict(rfModel, test)
ptest
```


# Create Files for Part 2 Submission
```{r}
answers <- as.vector(ptest)

pml_write_files = function(x) {
  n = length(x)
  for (i in 1:n) {
    filename = paste0("problem_id_", i, ".txt")
    write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
                col.names = FALSE)
  }
}

#pml_write_files(answers)

```



```

